---
title: "Intro to basics"
output:
  html_notebook:
    df_print: paged
  html_document:
    df_print: paged
description: In this chapter, you will take your first steps with R. You will learn
  how to use the console as a calculator and how to assign variables. You will also
  get to know the basic data types in R. Let's get started!
title_meta: Chapter 1
---

## Command prompt
Type commands on the line tat begins with a `>` sign (known as the prompt). When you hit enter, Rstudio will run your command and display any output below it
```{r}
1+1
```
R displays an index next to th output; **ignore it**. It`s helpful when R return more than one value in the output
```{r}
seq(300)
```

## R: the fancy calculator
```{r}
5+5
4-1
1*2
4^2
```


### Arithmetic with R

In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`

The last two might need some explaining:

- The `^` operator raises the number to its left to the power of the number to its right: for example `3^2` is `9`.
- The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or `5 %% 3` is 2.

With this knowledge, follow the instructions below to complete the exercise.

**Instructions**
- Type `2^5` in the editor to calculate 2 to the power 5.
```{r}
2^3
```
- Type `28 %% 6` to calculate 28 modulo 6.
```{r}
28 %% 6
```

- Note how the `#` symbol is used to add comments on the R code.

