---
title: "Intro to basics"
output:
  html_notebook:
    df_print: paged
  html_document:
    df_print: paged
---

## R Objetcs

### Variable Assignment
Save information as an R objetc with the *greater than* sign followed by *a minus*, e.g. an arrow: `<-`
```{r}
#name of new objetc         assignment operator, "gets"         information to store in the objetc
foo                                 <-                            42
```

### Common R Workflow
Save output of one function as an R objetc to use in a second function
```{r}
foo <- round(3.1415) + 1
foo
```

```{r}
factorial(foo)
```

## Remove objetcs from environment
You can remove an objetc with `rm`
```{r}
fac_foo<-factorial(foo)
fac_foo
```

```{r}
rm(foo)
rm(fac_foo)
```

### To consider...

- Object names cannot beggin with numbers. Wise to avoid names already in use.

```{r}
mean<-mean(rnorm(100))
mean
```
```{r}
rm(mean)
```

- R will treat the lowercase and capital letters as different.


## Data Structures

You can save more than a single number in an objetc by creating a *vector*, *matrix*, or *array*.

```{r}
WorldPhones
```


### Vectors

- Combine multiple elements into one dimensional array.

- Create with the `c` function.
```{r}
vec<-c(1,2,3,10,100)
vec
```

### Matrices

- Combine multiple elements into a two dimensional array.

- Create with the `matrix` function.
```{r}
            #vector of elements to go in the matrix     number of rows for matrix

mat<-matrix(data = c(1,2,3,4,5,6),                      nrow = 2)
mat
```

```{r}
            #vector of elements to go in the matrix     number of rows for matrix

mat<-matrix(data = c(1,2,3,4,5,6),                      nrow = 3)
mat
```


```{r}
            #vector of elements to go in the matrix     number of rows for matrix
#rbind(c(1,2,3,4,5,6),c(1,2,3,4,5,6)*2)
mat<-matrix(c(1,2,3,4,5,6),ncol=2)
mat
```

### Math: element- wise
```{r}
vec2<-vec
vec2<-vec + 4
vec; vec2
```

```{r}
vec * 4 ; vec2 * 4
```

```{r}
vec * vec ; vec2 * vec2; c(23,vec) * c(vec2,2);vec;vec2
```

### Matrix multiplication

**inner**
```{r}
mat<-matrix(c(1,2,3,4,5,6,7,8,9),ncol=3)
vec; vec %*% vec; mat;mat %*% mat
```

**outer**
```{r}

mat<-matrix(c(1,2,3,4,5,6),ncol=2)
vec; vec %o% vec; mat; mat %o% mat; mat %o% vec
```

**transpose**
```{r}
mat
```

```{r}
t(mat)
```

##Data types
R can recognize different types of data.

We'll look at four basic types:

- numbers
- character strings (text)
- logical
- factor


### numeric
- Any number, no quotes. 
- Appropiate for math.

```{r}
12+4
```

```{r}
3000000
```

```{r}
class(0.000001)
```

### character

- Any symbols surronded by quotes.
- Appropiate for words, variable names, messages, any text.

```{r}
"hello"
```

```{r}
class("hello")
```


```{r}
"12+4"
```

```{r}
class("12+4")
```

```{r}
"hello" + "world"
```

```{r}
nchar("hello")
```

```{r}
paste("hello","world",sep=",");paste("hello","world",sep=" 2342eafdsghsIJGBJmdxfghvb ")
```


```{r}
paste("hello","world",sep=",");paste("como","estas",sep="_");

paste(paste("hello","world",sep=","),paste("como","estas",sep="_"),2,sep="::"); paste(paste("hello","world",sep=","),paste("como","estas",sep="_"),"2",sep="::"); paste("hola","sin","espacios",sep="");paste0("hola","sin","espacios")
```

### To consider...
**wich are numbers?**

```{r}
1; "1"; "one"
```
```{r}
c(1, "1","one")
```
### logical
`TRUE` or `FALSE`
R's form of binary data. Useful for logical tests.

```{r}
3<4
```

```{r}
class(TRUE)
```

```{r}
class(T) ; class(F)
```

```{r}
class(3<4)
```

## factor
R's form of categorical data. Saved as an integer with a set of labels (e.g. levels)
```{r}
fac<-factor(c("a","b","c"))
fac
```
```{r}
class(fac)
```

### To consider...
```{r}
x<-c(1,2,3)
```

**what is the difference between these?**
```{r}
x; "x"
```
